---
title: "ESM 244 Lab 5"
author: "Sam Csik"
date: "2/7/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages and data
```{r}
library(tidyverse)
library(tseries)
library(forecast)

# US residential energy consumption values (Jan 19873 - Oct 2017), Trillion BTU (US Energy Information Administration)
energy <- read_csv("energy.csv")
```

Create 'ts' time series data using info in the res_total column from 'energy'
```{r}
# make time series class with this specified vector 
res_ts <- ts(energy$res_total, frequency = 12, start = c(1973, 1))

# exploratory plot of time series data
plot(res_ts)
```

Decompose ts
```{r}
res_dc <- decompose(res_ts)

plot(res_dc)
```

other cool ways to present ts data
```{r}
monthplot(res_ts) # breaks down by month how values are changing over all of the years that it's been recorded, plus the mean value
```

```{r}
ggseasonplot(res_ts, polar = TRUE)
```

Simple moving averages (changing the window drastically alters how the data appears--BE VERY CAREFUL...best to show original data + MA at the same time)
```{r}
sma_res <- ma(res_ts, order = 5)

plot(res_ts)
lines(sma_res, col = "red")

# can bind sma_res data to orignal df to finalize plot
```

ACF (autocorrelation)
```{r}
res_acf <- acf(res_ts)
ggtsdisplay(res_ts) # auto- and partial-correlation function
```

Holt-Winters exponential smoothing # gives more weight to more recent observations
```{r}
res_hw <- HoltWinters(res_ts)
plot(res_hw)
```

Future forecasting with Holt-Winters
```{r}
res_forecast <- forecast(res_hw, h = 60) # forecast for 5 years
plot(res_forecast)
```

Check distribution of residuals
```{r}
hist(res_forecast$residuals) # looks normally distributed!
```

Autoregressive integrated moving average (ARIMA) using auto.arima for p, d, q
```{r}
res_pdq <- auto.arima(res_ts) # first = seasonal component, second = nonseasonal component
```

pdq:[1,0,2][0,1,1]

Fit the ARIMA model using those pdq inputs:
```{r}
res_arima <- arima(res_ts, order = c(1,0,2), seasonal = list(order = c(0,1,1)))

# check out residuals
par(mfrow = c(1,2))
hist(res_arima$residuals)
qqnorm(res_arima$residuals)
```

Forecat with ARIMA outputs:
```{r}
par(mfrow = c(1,1))
forecast_res <- forecast(res_arima, h = 72)
plot(forecast_res)
```

View in ggplot
```{r}
res_df <- data.frame(forecast_res)
month_seq <- seq(1,72)

res_df_2 <- data.frame(month_seq, res_df)

ggplot(res_df_2, aes(x = month_seq, y = Point.Forecast)) +
  geom_line()
```

